---
title: "Experimental treatment with edaravone in a mouse model of spinocerebellar ataxia 1"
subtitle: "Data upload and defining functions"
author: "Martina Sucha, Simona Benediktova, Filip Tichanek, Jan Jedlicka, Stepan Kapl, Dana Jelinkova, Zdenka Purkartova, Jan Tuma, Jitka Kuncova, Jan Cendelin"
format: 
  html:
    embed-resources: true
    keep-md: true
    toc: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    number-depth: 4
    code-fold: true
    code-summary: "Open code"
editor: visual
project:
  type: default
  output-dir: output
theme: sandstone
fontsize: 13.5px
---

This page shows R code for the study [Sucha et al.](https://www.mdpi.com/1422-0067/24/13/10689) (2023, *International Journal of Molecular Science*).

**Citation**:

> Sucha M, Benediktova S, Tichanek F, Jedlicka J, Kapl S, Jelinkova D, Purkartova Z, Tuma J, Kuncova J, Cendelin J. Experimental Treatment with Edaravone in a Mouse Model of Spinocerebellar Ataxia 1. International Journal of Molecular Sciences. 2023; 24(13):10689. https://doi.org/10.3390/ijms241310689

[GitHub page](https://github.com/filip-tichanek/edaravonSCA1): https://github.com/filip-tichanek/edaravonSCA1

# Import of packages

```{r}
rm(list = ls())

suppressWarnings(suppressMessages( {
library(brms)
library(beeswarm)
library(vioplot)
library(glmmTMB)
library(car)
library(cowplot)
library(ggplot2)
library(tidyverse)
  } ) )
```

# Custom functions

## *run_model*
Function to run and save, or load, Bayesian model. The function was written according to online proposal of Paul-Christian Bürkner

```{r}
    run_model <- function(expr, path, reuse = TRUE) {
      path <- paste0(path, ".Rds")
      if (reuse) {
        fit <- suppressWarnings(try(readRDS(path), silent = TRUE))
        }
      if (is(fit, "try-error")) {
        fit <- eval(expr)
        saveRDS(fit, file = path)
        }
      fit
    }

```

## *prios*

The function takes model (***m***) fitted with *glmmTMB* package and show prior with sca1 as reference group and the WT genotype as the 1st b regression parameter. The model *m* is based on previous data ([Tichanek et al., 2020](https://www.nature.com/articles/s41598-020-62308-0), *Scientific Reports*).

```{r}
prios<- function(m){
  pr<-summary(m)$coefficients$cond[2,c(1,2,4)]
  if((pr[1]<0)&(pr[3]<0.05)){
    mu<-(pr[1]+(pr[2]*1.96))
    sigma<-max(abs(pr[1]-mu),abs(mu))*1.5}
  else if((pr[1]>0)&(pr[3]<0.05)){
    mu<-(pr[1]-pr[2]*1.96)
    sigma<-max(abs(pr[1]-mu),abs(mu))*1.5}
  else{
    mu=0
    sigma=0-abs(pr[1])*1.5}
  return(c(mu,sigma))}
```

## *mons_poste*

The function takes data frame of posterior distributions of effect sizes (***datf*** ) and value of zero effect (***tres***). There are two version: *mons_poste6* is for visualizing posterior probabilities on limited space when there are 6 posterior distribution in a single column. *mons_poste* is used otherwise.

```{r}
## mons_poste for visualisations of posterior distribution
mons_poste<-function(datf, tres){
  xx<-1;xax<-data.frame();yax<-data.frame();cisl<-data.frame()
  repeat{
    xax[1:512,xx] <-density(datf[,xx])$x
    yax[1:512,xx]<-density(datf[,xx])$y
    vec<-data.frame(datf[,xx])
    cisl[xx,1:7]<-sapply(vec,function(p) quantile(p,probs=c(0.25,0.75,0.025,0.975,0.001,0.999,0.5)))
    xx<-xx+1;if(xx>dim(datf)[2]){break}}
  
  axl<-data.frame(cbind(xax,yax))
  dim(axl)
  
  scaler<- max(xax) - min(xax)
  uniscal<-round(log10(160/scaler))
  
  xrange<- c(min(xax)-0.05*scaler,max(xax));xscaler=xrange[2]-xrange[1]
  zpos<-c(seq(0,1,by=1/dim(datf)[2]))
  ymax<-1/(max(yax)*1.8*dim(datf)[2])
  par(mgp=c(1.6,0.55,0))
  plot(NULL, xlim=c(xrange[1],xrange[2]),ylim=c(0,1),xlab="",ylab="",las=1, axes=FALSE)
  
  xxs<-1
  repeat{
   x<-ste[xxs]
   lines(c(x,x),c(0,1),col="grey70",lwd=0.7,lty=3)
    xxs<-xxs+1;if(xxs>length(ste)){break}}
  
  xx<-1
  linop=0.05
  repeat{
    polygon(xax[,xx],(ymax*yax[,xx])+(zpos[xx])+linop,col=rgb(1,0.9,0.7),border="grey50")
    polygon(xax[,xx],(ymax*yax[,xx])+(zpos[xx])+linop,col=rgb(0.82,0.82,0.82),border="NA")
    
    ax<- axl[,c(xx,xx+dim(datf)[2])] [axl[,xx]>cisl[xx,3]&axl[,xx]<cisl[xx,4],]
    polygon(c(ax[,1],rev(ax[,1])),
            c((ymax*ax[,2])+(zpos[xx])+linop,
              rep((zpos[xx])+linop,length(ax[,2]))
            ), col=rgb(1,0.88,0.46),border=NA)
    lines(c(cisl[xx,3],cisl[xx,4]),c((zpos[xx])+linop,(zpos[xx])+linop),lend=1,lwd=3,col="grey10")
    points(cisl[xx,7],zpos[xx]+linop,lend=2,pch=3,cex=2.2,lwd=1.5,col="grey10")
    
    
    xx<-xx+1
    if(xx>length(zpos)-1){break}}
  lines(c(tres,tres),c(0,1),lty=1,col="blue4")
  axis(side=1,las=1,cex.axis=1.1,
       at=c(ste),pos=range[1],tck=tckk)
  lines(c(min(xax)-scaler*0.05,max(xax)),c(range[1],range[1]))
  
  xx<-1
  repeat{
    lines(c(xrange[1]+0.2*xscaler,xrange[1]+1*xscaler),
          c(zpos[xx]+zpos[2]*0.88,zpos[xx]+zpos[2]*0.9),
          lwd=22,col="white",lend=2)
    text(xrange[1]+0.55*xscaler,zpos[xx]+zpos[2]*0.88,
         bquote(.(round(cisl[xx,7],uniscal)) *
                  " (" *.(round(cisl[xx,3],uniscal)) *
                  " to "* .(round(cisl[xx,4],uniscal)) * ")")
         ,xpd=TRUE,cex=1,col=rgb(0.4, 0.1, 0))
    pd<-p_dir(datf[,xx],tres)
    pd2<- -round(log10(1-pd))+1
    text(xrange[1]+0.9*xscaler,zpos[xx]+zpos[2]*0.6,
         bquote(.(round(pd,pd2)))
         ,xpd=TRUE,cex=1.1,col=rgb(0.6, 0.1, 0.8))
    xx=xx+1;if(xx>length(zpos)-1){break}}
  text(mean(c(xrange[1],max(xax))),range[1]-0.12,paste(yla),xpd=T,cex=1.1)
}


## mons_poste6 for visualisations of 6 posterior distributions in a single column of the figure 
mons_poste6<-function(datf, tres){
  xx<-1;xax<-data.frame();yax<-data.frame();cisl<-data.frame()
  repeat{
    xax[1:512,xx] <-density(datf[,xx])$x
    yax[1:512,xx]<-density(datf[,xx])$y
    vec<-data.frame(datf[,xx])
    cisl[xx,1:7]<-sapply(vec,function(p) quantile(p,probs=c(0.25,0.75,0.025,0.975,0.001,0.999,0.5)))
    xx<-xx+1;if(xx>dim(datf)[2]){break}}
  
  axl<-data.frame(cbind(xax,yax))
  dim(axl)
  
  scaler<- max(xax) - min(xax)
  uniscal<-round(log10(160/scaler))
  
  xrange<- c(min(xax)-0.05*scaler,max(xax));xscaler=xrange[2]-xrange[1]
  zpos<-c(seq(0,1,by=1/dim(datf)[2]))
  ymax<-1/(max(yax)*1.8*dim(datf)[2])
  par(mgp=c(1.6,0.55,0))
  plot(NULL, xlim=c(xrange[1],xrange[2]),ylim=c(0,1),xlab="",ylab="",las=1, axes=FALSE)
  
  xxs<-1
  repeat{
    x<-ste[xxs]
    lines(c(x,x),c(0,1),col="grey70",lwd=0.7,lty=3)
    xxs<-xxs+1;if(xxs>length(ste)){break}}
  
  xx<-1
  linop=0.01
  repeat{
    polygon(xax[,xx],(ymax*yax[,xx])+(zpos[xx])+linop,col=rgb(1,0.9,0.7),border="grey50")
    polygon(xax[,xx],(ymax*yax[,xx])+(zpos[xx])+linop,col=rgb(0.82,0.82,0.82),border="NA")
    
    ax<- axl[,c(xx,xx+dim(datf)[2])] [axl[,xx]>cisl[xx,3]&axl[,xx]<cisl[xx,4],]
    polygon(c(ax[,1],rev(ax[,1])),
            c((ymax*ax[,2])+(zpos[xx])+linop,
              rep((zpos[xx])+linop,length(ax[,2]))
            ), col=rgb(1,0.88,0.46),border=NA)
    lines(c(cisl[xx,3],cisl[xx,4]),c((zpos[xx])+linop,(zpos[xx])+linop),lend=1,lwd=3,col="grey10")
    points(cisl[xx,7],zpos[xx]+linop,lend=2,pch=3,cex=2.2,lwd=1.5,col="grey10")
    
    xx<-xx+1
    if(xx>length(zpos)-1){break}}
  lines(c(tres,tres),c(0,1),lty=1,col="blue4")
  axis(side=1,las=1,cex.axis=1.1,
       at=c(ste),pos=0,tck=tckk,xpd=TRUE)
  lines(c(min(xax)-scaler*0.05,max(xax)),c(range[1],range[1]))
  
  xx<-1
  repeat{
    lines(c(xrange[1]+0.2*xscaler,xrange[1]+1*xscaler),
          c(zpos[xx]+zpos[2]*0.88,zpos[xx]+zpos[2]*0.9),
          lwd=22,col="white",lend=2)
    text(xrange[1]+0.55*xscaler,zpos[xx]+zpos[2]*0.88,
         bquote(.(round(cisl[xx,7],uniscal)) *
                  " (" *.(round(cisl[xx,3],uniscal)) *
                  " to "* .(round(cisl[xx,4],uniscal)) * ")")
         ,xpd=TRUE,cex=1.2,col=rgb(0.4, 0.1, 0))
    pd<-p_dir(datf[,xx],tres)
    pd2<- -round(log10(1-pd))+1
    text(xrange[1]+0.9*xscaler,zpos[xx]+zpos[2]*0.6,
         bquote(.(round(pd,pd2)))
         ,xpd=TRUE,cex=1.2,col=rgb(0.6, 0.1, 0.8))
    xx=xx+1;if(xx>length(zpos)-1){break}}
  text(mean(c(xrange[1],max(xax))),0-0.12,paste(yla),xpd=T,cex=1.1)
}

```

## *p_dir*

Takes data frame of posterior distribution of effect size (***data***) and value of zero effect (***tres***) and calculates *probability of direction*

```{r}
p_dir<-function(data,tres){
  1-((min(length(data[data>tres]),length(data[data<tres])))/(length(data)))
}
```

## Functions for data transformations

```{r}
logit<-function(x){
  log(x/(1-x))
}

inv.logit<-function(x){
  exp(x)/(exp(x)+1)
  }

asinTransform <- function(p) { asin(sqrt(p)) }

revAsin <- function(p){sin(p)^2}
```

# Colors definitions

```{r}
alp=1;cola<-c(
  rgb(0.7,  0.7, 0.25,alpha=alp),
  rgb(0.55,0.55, 0.55,alpha=alp),
  rgb(0.9,0,0,alpha=alp),
  rgb(0.1, 0.1, 1,alpha=alp))

alp=0.2;colb<-c(
  rgb(0.7,  0.7, 0.25,alpha=alp),
  rgb(0.55,0.55, 0.55,alpha=alp),
  rgb(0.9,0,0,alpha=alp),
  rgb(0.1, 0.1, 1,alpha=alp))

alp=0.5;colc<-c(
  rgb(0.7,  0.7, 0.25,alpha=alp),
  rgb(0.55,0.55, 0.55,alpha=alp),
  rgb(0.9,0,0,alpha=alp),
  rgb(0.1, 0.1, 1,alpha=alp))

alp=0.4;cold<-c(
  rgb(0.7,  0.7, 0.25,alpha=alp),
  rgb(0.55,0.55, 0.55,alpha=alp),
  rgb(0.9,0,0,alpha=alp),
  rgb(0.1, 0.1, 1,alpha=alp))

cole=cola
```

# Data upload and wrangling

## Behavioral data

```{r}
## upload of data
behav<-read.csv("source_data/behav_dat.csv", stringsAsFactors = T)

## subject (id) as factor
behav$id <- as.factor(behav$id)

## defining factor 'group'
behav$group<-interaction(behav$treatment,behav$genotype)
behav$group<-factor(behav$group,levels=c('ctrl.wt','eda.wt','ctrl.sca','eda.sca'))

## defining numerical version of 'sex' variable
behav$sex_num<-ifelse(behav$sex=="f",-0.5,0.5)

## defining variable combining 'genotype', 'treatment' and 'sex' into single variable ('group_sex')
behav$group_sex<-interaction(behav$treatment,behav$genotype,behav$sex)
behav$group_sex<-factor(behav$group_sex, levels=c('ctrl.wt.f','ctrl.wt.m',
                                                 'eda.wt.f','eda.wt.m',
                                                 'ctrl.sca.f','ctrl.sca.m',
                                                 'eda.sca.f','eda.sca.m'))


## substetting data according to sex
behav_m<-subset(behav,behav$sex=="m")
behav_f<-subset(behav,behav$sex=="f")

## data summary
summary(behav)

```

## Rotarod data

```{r}
## upload
rotarod<-read.csv("source_data/rotarod_data.csv",stringsAsFactors = T)

## data summary
summary(rotarod)

```

## Data from [Tichanek et al., 2020, SciRep](https://www.nature.com/articles/s41598-020-62308-0)

```{r}
date_scirep<- read.csv("source_data/scirep_2020_data.csv",sep=";",stringsAsFactors = T)
young_scirep <- subset(date_scirep,date_scirep$age_cohort=="sca1_w06"|date_scirep$age_cohort=="sca1_w10")
young_scirep$genotype<-relevel(young_scirep$genotype,ref="tg+")

## data summary
summary(young_scirep)
```

## Mitochondrial respiration data

```{r}

## data upload 
mitochondria <- read.csv("source_data/mitochondria.csv", stringsAsFactors = T)

## data summary
summary(mitochondria)

```

## Calbindin immunofluorescence intensity

```{r}
## upload data of intensity in molecular layer
calb <- read.csv("source_data/calbindin.csv", sep = ';', header = TRUE, stringsAsFactors = T)

## subject as factor
calb$id <- as.factor(calb$id)

## slice as factor, nested in subject
calb$slice_id <- interaction(calb$id, calb$slice)

## defining groups
calb$group<-interaction(calb$treatment,calb$genotype)
calb$group<-factor(calb$group,levels=c('ctrl.wt','ctrl.sca','eda.sca'))

## sex as numerical variable, scale to -0.5 and 0.5 values
calb$sex_num<-ifelse(calb$sex=="f",-0.5,0.5)

## defining factor combining 'group' and 'sex'
calb$group_sex<-interaction(calb$treatment,calb$genotype,calb$sex)
calb$group_sex<-factor(calb$group_sex, levels=c('ctrl.wt.f','ctrl.wt.m',
                                                  'ctrl.sca.f','ctrl.sca.m',
                                                  'eda.sca.f','eda.sca.m'))

## upload of calbindin intensities in granular layer (GL)
calbgran <- read.csv("source_data/calbgran.csv", sep = ';', header = TRUE, stringsAsFactors = T)
calbgran$id <- as.factor(calbgran$id)
calbgran$slice_id <- interaction(calbgran$id, calbgran$slice)

## there are two GL measurements neighouring one molecular layer measurement. 
## Lets delete one to simplify combining of both datasets
calbgran <- calbgran %>% filter(slice_id != '13.9' | order != 'F')
calb <- calb %>% mutate(
  mean_gran = calbgran$mean)

calb$group <- factor(calb$group)

## data summary
summary(calb)

```

## PSA-NCAM data

```{r}

## upload data
psa <- read.csv('source_data/psa_ncam.csv', stringsAsFactors = T)

## upload data for connecting psa table with metadata
histid <- read.csv('source_data/hist_id.csv', stringsAsFactors = T)

## connecting both tables
psa <- left_join(psa, histid, by = 'hist_id')

## removing missing data
psa <- subset(psa, psa$psa_ncam != 'X' & psa$psa_ncam != '')
psa$psa_ncam <- as.numeric(as.character(psa$psa_ncam))

## defining psa_ncam and numerical variable
psa_nc <- subset(psa, grepl('control', as.character(psa$slice)) == FALSE)

## defning factor of slice order
psa_nc$slice <- factor(psa_nc$slice)

## slice order as numerical variable
psa_nc$slice_n <- as.numeric(as.character(psa_nc$slice))

## data summary
summary(psa_nc)

```

## Hippocampal volume data

```{r}
## upload data
hp_vol <- read.csv('source_data/hip_vol.csv', stringsAsFactors = T)

## defining 'group'
hp_vol$group<-interaction(hp_vol$treatment,hp_vol$genotype)
hp_vol$group<-factor(hp_vol$group,levels=c('ctrl.wt','ctrl.sca','eda.sca'))

## rescaling 'sex'
hp_vol$sex_num<-ifelse(hp_vol$sex=="f",-0.5,0.5)


## defining factor combining 'group' and 'sex'
hp_vol$group_sex<-interaction(hp_vol$treatment,hp_vol$genotype,hp_vol$sex)
hp_vol$group_sex<-factor(hp_vol$group_sex, levels=c('ctrl.wt.f','ctrl.wt.m',
                                                  'ctrl.sca.f','ctrl.sca.m',
                                                  'eda.sca.f','eda.sca.m'))

## data summary
summary(hp_vol)

```

## Cerebellar molecular layer volume data

```{r}
## data upload
cbml <- read.csv('source_data/cbml.csv', stringsAsFactors = T) 

## defining 'group'
cbml$group<-interaction(cbml$treatment,cbml$genotype)
cbml$group<-factor(cbml$group,levels=c('ctrl.wt','ctrl.sca','eda.sca'))

## rescaling 'sex' factor
cbml$sex_num<-ifelse(cbml$sex=="f",-0.5,0.5)

## defining 'group_sex'
cbml$group_sex<-interaction(cbml$treatment,cbml$genotype,cbml$sex)
cbml$group_sex<-factor(cbml$group_sex, levels=c('ctrl.wt.f','ctrl.wt.m',
                                                    'ctrl.sca.f','ctrl.sca.m',
                                                    'eda.sca.f','eda.sca.m'))
## subject as factor
cbml$id <- factor(cbml$id)

## rename the outcome variable
cbml$volume <- cbml$cb_ml

## data summary
summary(cbml)

```

## Citrate synthase activity data

```{r}

## data upload
csyn <- read.csv('source_data/citrate_synt.csv', stringsAsFactors = T)

## defining 'group'
csyn$group<-interaction(csyn$treatment,csyn$genotype)
csyn$group<-factor(csyn$group,levels=c('ctrl.wt','eda.wt','ctrl.sca','eda.sca'))

## subject as factor
csyn$id <- factor(csyn$id)

## data summary
summary(csyn)

```

# Saving the environment

Save the R environment to load data in other scripts

```{r}
if(T){save.image(file='source_data/myEnvironment.RData')} else {print('Env NOT SAVED NOW!')}
```

# Session info

```{r}
sessionInfo()
```
